Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-02-14 vesicle_code/docs/adhesion/adhesion.tex
changes as of Feb 12 9pm by yny

\title{Hydrodynamics and rheology of vesicle doublets}
% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

\begin{abstract}
The dynamics of an adhesive vesicle doublet under various flow conditions is
investigated numerically using a high-order, adaptive-in-time boundary
integral method. In a quiescent flow, two nearby vesicles
move slowly towards each other under the adhesive potential, pushing out
fluid between them to form a vesicle doublet at equilibrium. A lubrication
analysis on such draining of a thin film gives the dependencies of
draining time on adhesion strength and separation distance that are in
good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is dynamically
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the 
diverging stream, thus providing a simple method to measure the adhesion
strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form as two vesicles approach each other provided that the adhesion
strength is sufficiently large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of both reduced area and
the adhesion strength.  Results from these numerical studies and
analysis shed light on the hydrodynamic and rheological consequences of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering~\cite{sackmann1996, Barthes-Biesel2016_ARFM}.  The ever
expanding applications have encouraged detailed experimental,
theoretical, and numerical investigations of vesicle hydrodynamics in
applied flow, electric fields, and magnetic fields.  Theoretical
investigations often take the form of small-deformation analysis or
spheroidal dynamics of a nearly-spherical vesicle (or nearly-circular in
two dimensions) subjected to an applied flow or under an external
forcing~\cite{Barthes-BieselRallison1981_JFM, Misbah2006_PRL,
Vlahovska2007_PRE, Finken2008_EPL, ZhangZahnTanLin2013_PoF,
Nganguia2013_PRE}.  On the other hand, various numerical methods have
been developed for simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE, Biben2005_EJP,
Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.

@@

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or
clusters~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, an idealized model for adhesive vesicle-vesicle interactions
gives rise to vesicle configurations similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  Using the Lennard-Jones
(L.-J.)~potential for point-point interaction between two vesicle
membranes without the molecular details of adhesive interactions between
RBCs, Flormann {\em et al.}~\cite{FlormannAouane2017_SciReports} found
the contact zone to buckle under strong adhesion.  A direct consequence
of the buckling is the sigmoidal shape of the adhesion region, which is
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight into the
rheology of blood flow and how to use nano solutes to control the
hypertension by adjusting the adhesive interactions between RBCs.

@@

The electrostatic nature of lipid molecules (a hydrophobic tail and a
hydrophilic head with an electric dipole) complicates the interactions
between a lipid bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,
Book_IntermolecularSurfaceForces, PerutkovaFrank-Bertoncelij2013_CSB} or
a solid (such as a glass substrate or a nano particle).  Adhesion
between a vesicle and a substrate has been extensively studied with more
focus on the static equilibrium shapes~\cite{Seifert1990_PRA,
ShiFengGao2006_ActaMechSin, LinFreund2007_IntJSolidsStructures,
GruhnFrankeDimova2007_Langmuir,das2008adhesion, zhang2009phase,
SteinkuhlerAgudo-Canalejo2016_BJ} than the transient adhesion
process~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesive interactions between lipid membranes are essential in many
biomedical, biological, and biophysical processes.  For example, vesicle
adhesion is crucial to initiate membrane fusion and fission in
endocytosis and exocytosis and the transport of small vesicles through
membrane surfaces.  In the absence of an external electric field and
ions in the solvents, it is reported that the adhesive interactions
between two membranes can be well approximated by the L.-J.-type
potential~\cite{FlormannAouane2017_SciReports}, that consists of a
long-range attraction component and a short-range repulsion
component~\cite{Book_IntermolecularSurfaceForces}, and the combination
of the two gives rise to adhesion at the separation distance that
minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}.

@@

Finally, it is not clear how thermal fluctuations may affect the
hydrodynamics of vesicles under adhesion. For example, does the
fluctuating hydrodynamics in the thin film between two vesicles enhance
adhesion to keep vesicles bound under linear flows?  Recently Liu {\em
et al.}~\cite{LiuChuNewbyRead2018_bioRxiv} used immersed boundary
simulations to show that, at a separation distance of tens of
nanometers, the thin film between the two membranes facilitate the
coupling between membranes via strong hydrodynamic interactions. In
particular, they demonstrate numerically that the fluctuation in one
membrane is highly correlated to the other membrane without any physical
contact. We are actively pursuing this direction with hydrodynamic
modeling and simulations of adhesive membranes with thermal fluctions.

>>>>

\title{Hydrodynamics and rheology of a vesicle doublet suspension}
% Force line breaks with \\
%\thanks{A footnote to the article title}%

@@

\begin{abstract}
The dynamics of an adhesive vesicle doublet under various flow conditions is
investigated numerically using a high-order, adaptive-in-time boundary
integral method. In a quiescent flow, two nearby vesicles
move slowly towards each other under the adhesive potential, pushing out
fluid between them to form a vesicle doublet at equilibrium. A lubrication
analysis on such draining of a thin film gives the dependencies of
draining time on adhesion strength and separation distance that are in
good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is dynamically
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the 
diverging stream, thus providing a simple method to measure the adhesion
strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form provided that the adhesion
strength is sufficiently large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of  reduced area.  Results from these numerical studies and
analysis shed light on the hydrodynamic and rheological consequences of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering~\cite{sackmann1996, FenzSengupta2012_IntegrBiol,Barthes-Biesel2016_ARFM}.  The ever
expanding applications of vesicles have encouraged detailed experimental,
theoretical, and numerical investigations of vesicle dynamics in
applied flow, electric fields, and magnetic fields.  
Experimental investigations of vesicles have uncovered vesicle properties and various novel applications
\cite{Dobereiner2000_CurrentOpinionCIS,EvansRawiczSmith2013_FaradayDiscussions,SugiyamaToyota2018_Life}. 
Theoretical investigations are often limited to small-deformation analysis on a nearly spherical vesicle
or spheroidal analysis on a spheroidal vesicle \cite{Barthes-BieselRallison1981_JFM, Misbah2006_PRL,
Vlahovska2007_PRE, Finken2008_EPL, ZhangZahnTanLin2013_PoF,
Nganguia2013_PRE}.  On the other hand, various numerical methods have
been developed for simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE, Biben2005_EJP,
Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.

@@

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or
clusters~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, a simplified model for adhesive vesicle-vesicle interactions
can reproduce vesicle shapes similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  Using the Lennard-Jones
(L.-J.)~potential for point-point interaction between two vesicle
membranes without the molecular details of adhesive interactions between
RBCs, Flormann {\em et al.}~\cite{FlormannAouane2017_SciReports} found
that, under strong adhesion, 
the membrane may buckle to a sigmoidal shape in the contact region.
Such a sigmoidal vesicle shape is 
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight into the
rheology of blood flow and how to use nano solutes to control the
hypertension by adjusting the adhesive interactions between RBCs.

@@

The electrostatic nature of lipid molecules (a hydrophobic tail and a
hydrophilic head with an electric dipole) complicates the interactions
between a lipid bilayer membrane and another bilayer
membrane~\cite{EvansMetcalfe1984_BJ, Book_PhysicalBasisCellAdhesion,
Book_IntermolecularSurfaceForces, PerutkovaFrank-Bertoncelij2013_CSB} or
a solid (such as a glass substrate or a nano particle).  Adhesion
between a vesicle and a substrate has been extensively studied with more
focus on the static equilibrium shapes~\cite{Seifert1990_PRA,
ShiFengGao2006_ActaMechSin, LinFreund2007_IntJSolidsStructures,
GruhnFrankeDimova2007_Langmuir,das2008adhesion, zhang2009phase,Agudo-Canalejo2015_ACSNano,Agudo-CanalejoLipowsky2015_NanoLett,
SteinkuhlerAgudo-Canalejo2016_BJ,Agudo-CanalejoLipowsky2017_SoftMatt} than the transient adhesion
process~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesive interactions between lipid membranes are essential in many
biomedical, biological, and biophysical processes.  For example, vesicle
adhesion is crucial to initiate membrane fusion and fission in
endocytosis and exocytosis and the transport of small vesicles through
membrane surfaces.  In the absence of an external electric field and
ions in the solvents, it is reported that the adhesive interactions
between two membranes can be well approximated by the L.-J.-type
potential~\cite{FlormannAouane2017_SciReports}, that consists of a
long-range attraction component and a short-range repulsion
component~\cite{Book_IntermolecularSurfaceForces}, and the combination
of the two gives rise to adhesion at the separation distance that
minimizes the interaction
potential~\cite{Book_IntermolecularSurfaceForces}.

@@

Finally, it is not clear how thermal fluctuations may affect the
hydrodynamics of vesicles under adhesion. For example, does the
fluctuating hydrodynamics in the thin film between two vesicles enhance
adhesion to keep vesicles bound under linear flows as speculated in cells \cite{FenzBihrSchmidt2017_NaturePhys}?  Recently Liu {\em
et al.}~\cite{LiuChuNewbyRead2018_bioRxiv} used immersed boundary
simulations to show that, at a separation distance of tens of
nanometers, the thin film between the two membranes facilitate the
coupling between membranes via strong hydrodynamic interactions. In
particular, they demonstrate numerically that the fluctuation in one
membrane is highly correlated to the other membrane without any physical
contact. We are actively pursuing this direction with hydrodynamic
modeling and simulations of adhesive membranes with thermal fluctuations.
